---
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r global_opts, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE)
```
# Statistical Inference Part - 1
#### *Niraj Nair*
#### *6 June, 2020*

## Overview
#### A pdf report is created by answering the questions in the problem statement. This report consists of graphs and calculations to answer the questions. The pdf report is made using knitr. The main tasks done here are simualtions to explore inference and simple inferential data analysis.

## Problem Statement
#### In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

#### Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

#### 1. Show the sample mean and compare it to the theoretical mean of the distribution.
#### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
#### 3. Show that the distribution is approximately normal.

#### In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

## Preparation

#### Loading require packages.

```{r prep}
library(ggplot2)
set.seed(1)
```
#### Defining the variables. Sample Size = 40, Mean = .2, Number of simulations = 1000.
```{r defVar}
n <- 40
lambda <- .2
sim <- 1000
```
#### Creating the dataset using rexp.
```{r dataSet}
dataset <- matrix(rexp(n*sim, lambda), sim)
dim(dataset)
```
## Answers 
#### 1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
#### __The theoretical mean is 1 / lambda ie. 1/0.2 = 5.__
```{r mean}
theoMean <- 1 / lambda
sampleMeans <- apply(dataset, 1, mean)
calMean <- mean(sampleMeans)
round(calMean,2)
```
#### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
#### The theoretical SD is (1 / lambda) * (1/sqrt(n)) Therefore, theoretical variance is square of theroretical SD.
```{r variance}
theoSD <- (1 / lambda) * (1/sqrt(n))
theoVar <- theoSD^2
round(theoSD, 3)#Theoretical SD
round(theoVar, 3) #Theoretical Variance
calVar <- var(sampleMeans)
calSD <- sd(sampleMeans)
round(calSD, 3) #Calculated SD
round(calVar,3) #Calculated Variance
```
#### 3.Show that the distribution is approximately normal.
```{r normal}
sampleMeansData <- data.frame(sampleMeans)
g <- ggplot(data = sampleMeansData, aes(x=sampleMeans)) 
g + geom_histogram(binwidth = lambda, fill = "white", color="black", aes(y = ..density..), show.legend = TRUE, legendPosition = "top") +
    geom_vline(xintercept = theoMean, color = "yellow", size = 1) +
    geom_vline(xintercept = calMean, color = "red", size = .8, linetype = "dashed") +
    stat_function(fun = dnorm, args = list(mean = theoMean, sd = theoSD), aes(color = "green"), size = 1) +
    stat_function(fun = dnorm, args = list(mean = calMean, sd = calSD), aes(color = "magenta"), size = 1) + 
    labs(title = "Exponential Distribution", x = "Sample Means", y = "Density")+
    scale_colour_identity(name = "Normal Distribution",labels = c("Theoretical", "Calculated"),guide = "legend")
```

#### Here,
#### 1. White line -> Theoretical Mean
#### 2. Red dashed line-> Calculated mean

## Conclusion

#### 1. It is clear that that the center of distribution is approximately equal to the theoretical mean.
#### 2. The Standard deviation and Variance of the distribution is compared to their theoretical values and it is found that they are approximately equal.
#### 3. It is clear from the figure that the density distribution forms a bell curve and since the theoretical and calculated density distribution are very close to overlapping we can say that this is a Normal Distribution.































