---
output:
  pdf_document: default
  html_document: default
---
```{r global_opts, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```
# Statistical Inference Part - 2
# Basic inferential data analysis
#### *Niraj Nair*
#### *6 June, 2020*

## Assignment
#### Now in the second portion of the project, we’re going to analyze the ToothGrowth data in the R datasets package.
#### 1. Load the ToothGrowth data and perform some basic exploratory data analyses
#### 2. Provide a basic summary of the data.
#### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there’s other approaches worth considering)
#### 4. State your conclusions and the assumptions needed for your conclusions.
***

## Required Packages
```{r pack}
library(ggplot2)
library(dplyr)
```
## 1. Loading the ToothGrowth data and performing some basic exploratory data analyses.
```{r load}
data("ToothGrowth")
str(ToothGrowth)
head(ToothGrowth)
```

## 2. Obtaining a summary of the dataset.
```{r summary}
summary(ToothGrowth)
g <- ggplot(data = ToothGrowth, aes(x = dose, y = len, fill = factor(dose)))
g + geom_boxplot()+
    facet_grid(.~supp)+
    labs(title = "Tooth Growth from different supplements", x = "Dose", y = "Tooth Length") +
    scale_fill_discrete(name = "Dosage")
growth <- ToothGrowth %>% group_by(supp, dose) %>% summarise(len = mean(len))
growth
```
## 3. Using confidence intervals and hypothesis tests to compare tooth growth by supp and dose.
#### (a) We assume the null hypothesis to be that there is no relation betweem tooth growth and supp.
```{r htest}
OJ <- ToothGrowth$len[ToothGrowth$supp == "OJ"]
VC <- ToothGrowth$len[ToothGrowth$supp == "VC"]
t.test(OJ, VC, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = .95)
```
#### Since the P-value is 3% ie. < 5%, we can reject the null hypothesis. This conveys that the alternative hypohtesis is accepted and therefore there exists a relation between tooth lenght and supp.
#### (b) We assume the null hypothesis to be that there is no relation betweem tooth growth and dose.
```{r dosePrep}
d05 <- ToothGrowth$len[ToothGrowth$dose == .5]
d1 <- ToothGrowth$len[ToothGrowth$dose == 1]
d2 <- ToothGrowth$len[ToothGrowth$dose == 2]
```
* First we will check that if the alternative hypothesis is right for the smaller doses
```{r doseSmall}
t.test(d05, d1, alternative = "less", paired = FALSE, var.equal = FALSE, conf.level = .95)
```
* Second we will check that if the alternative hypothesis is right for the bigger doses
```{r doseBig}
t.test(d1, d2, alternative = "less", paired = FALSE, var.equal = FALSE, conf.level = .95)
```
#### From both the tests we can clearly see that the P-values are very low. Therefore the null hypothesis is rejected and we can say that the tooth growth depends on the doasage of supp.

## Conclusion
* There doesn't seem to be a statistically significant difference between delivery methods, with OJ apparently more effective at dose levels 0.5 and 1, and VC slightly more effective at dose level 2.

* It appears that there is a statistically significant difference between tooth length and dose levels across both delivery methods, in other words, as the 
dose increases so does tooth length.

















```














